Efficient FPGA Implementation of\\ Digit Parallel Online Arithmetic Operators

### 2014-05-21 FPT2014/bare_conf.tex
Self check 2

\begin{abstract}
%\boldmath
Online arithmetic has been widely utilized for ASIC implementations with significant performance improvements, as it is designed to perform computations from the most significant digits. Recent research also shows that the digit parallel online operators can fail more gracefully when operating beyond the deterministic region, in comparison to the operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators requires large area overhead, and the efficient implementation of general purpose online operators on FPGAs remains under investigation. In this paper, we propose novel approaches to implement the online operators such as adder and multiplier efficiently on current FPGAs with 6-input LUTs. The FPGA carry resources are used for both area reduction and performance improvements. We demonstrate experimentally that the proposed architectures achieve significant area savings and frequency speed-ups comparing to the original implementations.

@@

\section{Introduction}\label{Sec:Intro}
In the conventional form of computer arithmetic, the computation results are generated either from the least significant digit (LSD), e.g. addition and multiplication, or from the most significant digit (MSD), e.g. division and square root. This inconsistency in computing directions will potentially result in large latency when propagating data among different operations. Online arithmetic was designed to solve this problem \cite{Ercegovac_OnlineOverview,Ercegovac_Book}. With online arithmetic, both inputs and outputs are processed in a MSD-first manner. This enables parallelism and duplication among various operations, and the overall computation latency can be significantly reduced [xxx]. A brief overview of online arithmetic is given in Section~\ref{Sec:Background}.

Recent research has demonstrated another key advantage of online arithmetic that it allows gracefully degradation when operating beyond the deterministic region, e.g. under overclocking. This stream of research stems from the inspiration that for certain applications, releasing the absolute accuracy requirements could lead to significant performance improvements. However, traditional form of computer arithmetic does not fail gracefully when timing violation happens, because timing errors initially affect MSDs of the results. In comparison, online arithmetic is more ``overclocking friendly'', as timing errors only occurs at LSDs of the results. Despite its attractiveness, the usage of online arithmetic on existing FPGAs is still limited due to the huge area overhead.

Generally, it is difficult to efficiently map the arithmetic circuits solely onto the look-up-tables (LUTs), which are fundamental building blocks of FPGAs. Lots of efforts has been done to alleviate this concern. For instance, both main commercial FPGA vendors introduced dedicated carry logic into the logic block to reduce the carry propagation delay by over one order of magnitude~\cite{Virtex6}. There is a large amount of studies exploring methods that can effectively map circuits using the carry logic for performance improvements~\cite{FPL10_FPGA_CarryChain,Prefix_CarryChain}, or propose alternative carry logic for FPGAs~\cite{FPGA_CarryChain_New1,FPGA_CarryChain_New2} . Besides, hard DSP blocks are also included to boost multiplications and multiply-accumulations [xxx]. However, most of these approaches are designed to accelerate computations with the conventional arithmetic.

For online arithmetic, we notice from existing literatures that there are 2 directions that are worth investigating. First, although there has been previous research about FPGA implementations of online arithmetic, they typically only focus on online adders. A brief review of relevant works in this area is presented in Section~\ref{Sec:OA_FPGA}. However, the implementation of other key arithmetic primitives, such as online multipliers, are lack of exploration. Second, we found that existing approaches of building general purpose online adders are specifically for FPGAs with 4-input LUTs, and 2 LUTs within a Slice. However, as will be discussed in Section~\ref{Sec:OA_FPGA}, these approaches cannot be directly applied on many current FPGAs with 6-input LUTs and 4 LUTs in a slice, such as the Xilinx Virtex series FPGAs and all Xilinx 7 series FPGAs~\cite{Virtex7}.

In this paper, both issues are addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach of mapping the digit parallel online adder onto FPGA with 6-LUTs. The mapping is based on the fast carry logic, and the available resources within a slice can be fully utilized. In addition, in Section~\ref{Sec:OM_FPGA} we optimize the online multiplication algorithm to yield an efficient FPGA implementation of digit parallel online multiplier. The theoretical resource usage of both operators is given, and it is verified experimentally on Xilinx Virtex-6 FPGAs. We demonstrate that the proposed designs achieves significant area reduction and performance gain over their original implementations.

The main contributions of this paper are as follows:
\begin{enumerate}
    \item An efficient approach of mapping online adders onto modern FPGAs;
    \item Area efficient implementation of digit parallel online multiplication algorithm on FPGAs;
\end{enumerate}

@@

Online arithmetic has been widely used in numerous applications such as xxx [xxx]. Online arithmetic was originally designed for digit-serial operation, of which the data flow is illustrated in Fig~\ref{Fig:OnlineDataFlow}. It can be seen that in order to generate the first output digit, $\delta$ digits of inputs are required and $\delta$ is called ``online delay''. Notice that $\delta$ is normally a constant, which is independent of the precision in a given operation. For ease of discussion, in the following of this paper the input data are normalized to fixed point numbers in the range $(-1,1)$. Based on this premise, the online representation of $N$-digit operands and result at iteration $j$ are given by (\ref{Eq:Online_Operands}), where $j\in[-\delta,N-1]$ and $r$ denotes the radix \cite{Ercegovac_Book}.

\begin{figure}[tbp]
  \centering
  %\vspace{-2.5ex}
  \includegraphics[width=.48\textwidth]{./Figures/OnlineArithmetic_DataFlow.pdf}
  %\vspace{-3ex}
  \caption{Dataflow in digit-serial online arithmetic, in which both inputs and outputs are processed from the MSD to the LSD. $\delta$ denotes the online delay.}
  %\vspace{-.5ex}
  \label{Fig:OnlineDataFlow}
\end{figure}

\begin{eqnarray}\label{Eq:Online_Operands}
\footnotesize
  X_{[j]}=\sum_{i=1}^{j+\delta}x_ir^{-i},~Y_{[j]}=\sum_{i=1}^{j+\delta}y_ir^{-i},~Z_{[j]}=\sum_{i=1}^{j}z_ir^{-i}
\normalsize
\end{eqnarray}

MSD-first operation is possible with the employment of the redundant number system . Normally there are 2 most commonly used redundant number representation: carry-save (CS) \cite{CSadder} and signed-digit (SD) \cite{RedundantNumber}. For instance with SD representation, each digit is represented with a redundant digit set $\{-a, \cdots,-1,0, 1, \cdots, a\}$, where $a\in[r/2,r-1]$. In comparison, the standard non-redundant representation only uses a digit set $\{0,r-1\}$. Thus a standard number corresponds to several possible redundant representations. For example, the two's complement number $0.111$ can be represented in the online form as $0.10\overline{1}$, $0.0\overline{1}1$ and $0.111$ among many other possible representations.

@@

\subsection{Binary Online Multiplication}
Multiplication is another key arithmetic operator. Typically the online multiplication is performed in a recursive digit-serial manner, as illustrated in Algorithm~\ref{Algorithm:OnlineMult} \cite{Ercegovac_OnlineMult} where both inputs and outputs are $N$-digit number as given in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated with MSD-first through a selection function $sel()$. For the radix $r$ and a chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence \cite{Ercegovac_OnlineMult}. As radix-2 is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case $sel()$ is given by (\ref{Eq:SelFunc_OM}) \cite{Oregon_OnlineNetwork}. Notice that the election is made only based on 1 integer bit and 1 fractional bit of $W_{[j]}$.

@@

It is worth noting that the precision of the product $Z$ can be determined by directly truncating from the LSDs, as there is no carry propagations from the LSD to the MSD. This is useful especially when the multiplication product is used by successive computations, where only the first $N$ digits of the results are normally used. In comparison, the product is generated starting from the LSD with conventional arithmetic. Hence the successive operations cannot commence until the full-precision result is generated. This feature of online arithmetic has inspired research about dynamically control of computation precision in numerous applications.

@@

Specifically in type 1, both works took advantage of the built-in carry resources in FPGAs. Conventionally the ASIC implementation of online adders is based on the 4:2 compressors, as shown within the gray background in Fig.~\ref{Fig:PreviousWork}. However, directly applying this approach in the FPGAs could be less efficient. This is because there is no carry propagation between the 2 full adders within a 4:2 compressor, and the net delay between them can be large. Instead, Kamp et al and Ortiz et al described very similar mapping techniques for online adders with 2 different data representations, respectively. The main idea is to map the logic block within the dotted circle in Fig.~\ref{Fig:PreviousWork}. In this case, the fast-carry logic in the FPGA can be employed, and the delay between the 2 FAs is largely reduced.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder_42comp_Spartan.pdf}
	\caption{Map the online adder onto Spartan FPGAs using the fast-carry resources. The grey background highlights the 4:2 compressor. Dotted circle indicates the logic block (LB) which can be mapped to the FPGA carry resources.}
	\label{Fig:PreviousWork}
\end{figure}

However, we notice the major limitations of both approaches that they only target on FPGAs with 4-input LUTs and 2 LUTs within a logic slice, such as the Xilinx Spartan series and the Altera Cyclone series. This is naturally reasonable because one LB can be mapped to a single slice. Nevertheless, for FPGAs with 6-input LUTs (6-LUT) and 4 LUTs in a slice, such as the Xilinx Virtex series and all Xilinx 7 series FPGAs, directly applying the approaches in \cite{FPT09RA} or \cite{ASAP09RA} will result in either resource waste or logic fault. For instance, if two logic blocks are mapped to a slice with 4 LUTs as seen in Fig.~\ref{Fig:PreWorkVirtex6}, the outputs of \texttt{LB1} will be faulty because the its carry input cannot be explicitly initialized.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/SDadder_FastCarry_PreviousWork.pdf}
	\caption{An example illustrating that direct applying approaches in \cite{FPT09RA} on a Virtex-6 FPGA will results in faulty outputs, because the carry input of \texttt{LB1} cannot be explicitly initialized.}
	\label{Fig:PreWorkVirtex6}
\end{figure}

\subsection{Proposed Mapping Method}
To tackle this problem, we first modify the structure of the online adder to enable an efficient FPGA mapping. The structure of a 4-digit online adder is given as an example in Fig.~\ref{Fig:ModifiedOA}. In this equivalent structure, the first FA in each 4:2 compressor is split into 2 parts, which only generate carry and sum respectively. In this case they can be mapped individually on 2 LUTs. The second FA, which generates the outputs, is unchanged and can be implemented using the fast-carry logic.

@@

\begin{figure*}[tbp]
	\centering
	\includegraphics[width=.85\textwidth]{./Figures/SDadder_FPGA_New_4digit.pdf}
	\caption{Modified structure of online adder. Left: an example of 4-digit online adder. The shaded part refers to the 2 logic blocks (LBs) that can be mapped onto 1 slice. Right: one LB. the dotted box outlines the logic that can be mapped onto 1 LUT and the corresponding fast-carry logic.}
	\label{Fig:ModifiedOA}
\end{figure*}

The detailed slice mapping of the 2 LBs is shown in Fig.~\ref{Fig:SliceNew}. The I/O signals are identical to the previous example in Fig.~\ref{Fig:ModifiedOA}. It can be seen that a 6-LUT can be configured with two different output ports O6 and O5. For LB1, the carry input can be initialized by setting the O6 of LUT2 equal to 0 constantly. In this case, the output of MUXCY2 is O5 of LUT2, and the carry from LB2 will not affect the results of LB1. Using this mapping method the resources within 1 slice can be fully utilized, potentially leads to significant area reduction.

@@

\subsection{Performance Analysis}\label{subsec:OA_Performance}
In our experiments, we compare the proposed online adder (\texttt{OA\_new}) against the original design which is implemented using behavioural descriptions based on 4:2 compressors (\texttt{OA\_behv}). Notice that in order to utilize the carry resources and to avoid logic optimizations by the synthesis tool, the Xilinx primitive component "\texttt{Carry4}" is used to create \texttt{OA\_new}. In addition, we also include ripple carry adder (\texttt{RCA}) which uses conventional arithmetic, into the comparisons. All designs are targeting on Xilinx Virtex-6 FPGA with speed grade -1.

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.1. It can be seen that both designs with online arithmetic achieves relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay in the online adder is ideally a fixed value and it is irrelevant to the operand precisions. In comparison, the rated frequency of \texttt{RCA} drops drastically for larger designs. In this case, both online adders will eventually outperform \texttt{RCA}. However our approach operates constantly faster than \texttt{OA\_behv} for any operand word-length. The speed-ups of \texttt{OA\_new} over \texttt{OA\_behv} are from $20\%$ to $26\%$.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/Exp/Adder_Freq.eps}
	\caption{Rated frequencies of the RCA and the online adder with different implementation methods. The results are obtained from post place and route timing reports in ISE 14.1.}
	\label{Fig:AdderFreq}
\end{figure}

We compare the resource usages of all three adder designs with respect to a variety of operand word-lengths. The results are shown in Fig.\ref{Fig:AdderArea}. The theoretical area budgets of the online adder predicted by (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}) are also plotted with dotted lines in Fig.\ref{Fig:AdderArea}. Several observations can be made from the results. Firstly the predicted resource usages of both LUTs and Slices match well the results obtained from real implementations. Secondly it can be seen that the behavioural implementation \texttt{OA\_behv} is not area efficient for FPGAs, because the dedicated FPGA resources are not fully utilized. In comparison, our implementation \texttt{OA\_new} achieves significant area savings: $25\%\sim 33\%$ in LUTs and $67\%\sim77\%$ in Slices. Thirdly, our design still costs more area than \texttt{RCA}, with the area overhead of $80\%\sim98\%$ in terms of LUTs, and $0\%\sim88\%$ in terms of Slices. This is because for a single adder, the area overhead brought by the redundant number representations cannot be removed. However research has shown that this overhead can be reduced for multiple-operand adders or adder trees [xxx].

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_LUTs2.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_Slices2.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different binary adder implementations with respect to a variety of operand word-lengths. (a) Usage of LUTs. (b) Usage of Slices.}
  \label{Fig:AdderArea}
  %\vspace{-2ex}
\end{figure}

\section{Digit Parallel Online Multiplier on FPGAs}\label{Sec:OM_FPGA}
\subsection{Algorithm of Digit Parallel Online Multiplication}
Algorithm \ref{Algorithm:OnlineMult} as described in Section~\ref{Sec:Background} can be synthesized into a unrolled digit parallel structure. Recent study has shown that the digit-parallel online multiplier is more tolerant to timing violations in comparison to conventional arithmetic, because timing errors initially affect the least significant digits and this will lead to small error magnitude[xxx]. This implementation method is straightforward and can be utilized to design other digit parallel online operators.

On the other hand, significant area budget is required in order to implement this type of digit parallel online multiplier (\texttt{OM}). Typically this is due to three main reasons. First, in comparison to the conventional multiplier, more iteration stages are needed. For instance in an $N$-digit \texttt{OM}, totally $(2N+\delta)$ iterations are required to generate $2N$ digits outputs, and $(N+\delta)$ iterations for the most significant $N$ digits result. Second, in each stage the logic is complex as it involves operations such as online addition, digit-vector multiplication, shifting and other combination logic such as the selection function. Besides, the word-length of signals $H_{[j]}$, $W_{[j]}$ and $P_{[j]}$ is $N+2+\delta$, because the selection function needs 2 integer digits as well as 1 fraction digits to generate $z_j$, as described in Section~\ref{Sec:Background}. Whereas in an $N$-digit array multiplier, there are only $(N-1)$ rows, each of which is basically an $N$-digit ripple-carry adder. Third, this \texttt{OM} architecture is not optimized specifically for FPGA technology.

>>>>

\begin{abstract}
%\boldmath
Online arithmetic has been widely utilized for ASIC implementations with significant performance improvements, as it is designed to perform computations from the most significant digits. Recent research also shows that the digit parallel online operators can fail more gracefully when operating beyond the deterministic region, in comparison to the operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators requires large area overhead, and the efficient implementation of general purpose online operators on FPGAs remains under investigation. In this paper, we propose novel approaches to implement key primitives of online arithmetic, such as adders and multipliers, efficiently on modern FPGAs with 6-input LUTs. The FPGA carry resources are used for both area reduction and performance improvements. We demonstrate experimentally that the proposed architectures achieve significant area savings and frequency speed-ups comparing to the original implementations.

@@

\section{Introduction}\label{Sec:Intro}
In the conventional form of computer arithmetic, the computation results are generated either from the least significant digit (LSD), e.g. addition and multiplication, or from the most significant digit (MSD), e.g. division and square root. This inconsistency in computing directions will potentially result in large latency when propagating data among different operations. Online arithmetic was designed to solve this problem \cite{Ercegovac_OnlineOverview,Ercegovac_Book}. With online arithmetic, both inputs and outputs are processed in a MSD-first manner. This enables parallelism and duplication among multiple operations, and the overall computation latency can be significantly reduced. A brief overview of online arithmetic is given in Section~\ref{Sec:Background}.

Recent research has demonstrated another key advantage of online arithmetic that it allows gracefully degradation when operating beyond the deterministic region, e.g. under overclocking \cite{SKDAC14}. This stream of research stems from the inspiration that for certain applications, releasing the absolute accuracy requirements could lead to significant performance improvements. However, traditional form of computer arithmetic does not fail gracefully when timing violation happens, because timing errors initially affect MSDs of the results. In comparison, online arithmetic is more ``overclocking friendly'', as timing errors only occurs at LSDs of the results.

Despite its attractiveness, the usage of online arithmetic on existing FPGAs is still limited due to the huge area overhead. Generally, it is difficult to efficiently map the arithmetic circuits solely onto the look-up-tables (LUTs), which are fundamental building blocks of FPGAs. Lots of efforts has been done to alleviate this concern. For instance, both main commercial FPGA vendors introduced dedicated carry logic into the logic block to reduce the carry propagation delay by over one order of magnitude~\cite{Virtex6}. There is a large amount of studies exploring methods that can effectively map circuits using the carry logic for performance improvements~\cite{FPL10_FPGA_CarryChain,Prefix_CarryChain}, or propose alternative carry logic for FPGAs~\cite{FPGA_CarryChain_New1,FPGA_CarryChain_New2} . Besides, hard DSP blocks are also included to boost multiplications and multiply-accumulations~\cite{XilinxDSP}. However, most of these approaches are designed to accelerate computations with the conventional form of arithmetic.

For online arithmetic, we notice from existing literatures that there are 2 directions that are worth investigating. First, although there has been previous research about FPGA implementations of online arithmetic, they typically only focus on online adders. A brief review of relevant works in this area is presented in Section~\ref{Sec:OA_FPGA}. However, the implementation of other key arithmetic primitives, such as online multipliers, are lack of exploration. Second, we found that existing approaches of building general purpose online adders are specifically for low cost FPGAs with 4-input LUTs and 2 LUTs within a Slice, such as the Altera Cyclone III \cite{AlteraCyclone} and Xilinx Spartan 3 series \cite{XilinxSpartan}. However, as will be discussed in Section~\ref{Sec:OA_FPGA}, these approaches cannot be directly applied on many modern FPGAs with 6-input LUTs and 4 LUTs in a slice, such as the Xilinx Virtex series FPGAs and all Xilinx 7 series FPGAs~\cite{Virtex7}.

In this paper, both issues will be addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach of mapping the digit parallel online adder onto FPGA with 6-LUTs. The mapping is based on the fast carry logic, and the available resources within a slice can be fully utilized. In Section~\ref{Sec:OM_FPGA}, we optimize the online multiplication algorithm to yield an efficient FPGA implementation of digit parallel online multiplier. In addition, the theoretical resource usage of both operators is given, and they are verified experimentally on Xilinx Virtex-6 FPGAs. We demonstrate that the proposed designs achieve significant area reduction and performance gain over their original implementations. In summary, the main contributions of this paper are as follows:
\begin{enumerate}
    \item An efficient approach of mapping online adders onto modern FPGAs;
    \item Area efficient implementation of digit parallel online multiplication algorithm on FPGAs;
    \item Theoretical area models of the proposed architectures.
\end{enumerate}

@@

Online arithmetic has been widely used in numerous applications such as signal processing and control algorithms~\cite{Online_FPGADSP,Online_Control}. Online arithmetic was originally designed for digit-serial operation, of which the data flow is illustrated in Fig~\ref{Fig:OnlineDataFlow}. It can be seen that in order to generate the first output digit, $\delta$ digits of inputs are required and $\delta$ is called ``online delay''. Notice that $\delta$ is normally a small constant, which is independent of the precision in a given operation. For ease of discussion, in the following of this paper the input data are normalized to fixed point numbers in the range $(-1,1)$. Based on this premise, the online representation of $N$-digit operands and result at iteration $j$ are given by (\ref{Eq:Online_Operands}), where $j\in[-\delta,N-1]$ and $r$ denotes the radix \cite{Ercegovac_Book}.
%
\begin{figure}[tbp]
  \centering
  %\vspace{-2.5ex}
  \includegraphics[width=.48\textwidth]{./Figures/OnlineArithmetic_DataFlow.pdf}
  %\vspace{-3ex}
  \caption{Dataflow in digit-serial online arithmetic, in which both inputs and outputs are processed from the MSD to the LSD. $\delta$ denotes the online delay.}
  %\vspace{-.5ex}
  \label{Fig:OnlineDataFlow}
\end{figure}
%
\begin{eqnarray}\label{Eq:Online_Operands}
\footnotesize
  X_{[j]}=\sum_{i=1}^{j+\delta}x_ir^{-i},~Y_{[j]}=\sum_{i=1}^{j+\delta}y_ir^{-i},~Z_{[j]}=\sum_{i=1}^{j}z_ir^{-i}
\normalsize
\end{eqnarray}

MSD-first operation is possible with the employment of the redundant number system. Normally there are 2 most commonly used redundant number representation: carry-save (CS) \cite{CSadder} and signed-digit (SD) \cite{RedundantNumber}. For instance with SD representation, each digit is represented using a redundant digit set $\{-a, \cdots,-1,0, 1, \cdots, a\}$, where $a\in[r/2,r-1]$. In comparison, the standard non-redundant representation only uses a digit set $\{0,r-1\}$. Thus a standard number corresponds to several possible redundant representations. For example, the two's complement number $0.111$ can be represented in the online form as $0.10\overline{1}$, $0.0\overline{1}1$ and $0.111$ among many other possible representations.

@@

\subsection{Binary Online Multiplication}
Multiplication is another key arithmetic operator. Typically the online multiplication is performed in a recursive digit-serial manner, as illustrated in Algorithm~\ref{Algorithm:OnlineMult} \cite{Ercegovac_OnlineMult} where both inputs and outputs are $N$-digit number as given in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated with MSD-first through a selection function $sel()$. For the radix $r$ and a chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence \cite{Ercegovac_OnlineMult}. As radix-2 is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case $sel()$ is given by (\ref{Eq:SelFunc_OM}) \cite{Oregon_OnlineNetwork}. Notice that the election is made only based on 2 integer digits and 1 fractional digit of $W_{[j]}$.

@@

It is worth noting that the precision of the product $Z$ can be determined by directly truncating from the LSDs, since there is no carry propagations from the LSD to the MSD. This is useful especially when the multiplication product is used by successive computations, where only the first $N$ digits of the results are normally used. In comparison, the product is generated starting from the LSD with conventional arithmetic. Hence the successive operations cannot commence until the result with full-precision is generated. This feature of online arithmetic has inspired research about dynamically control of computation precision in numerous applications~\cite{Online_Truncation}.

@@

Specifically in type 1, both works took advantage of the built-in carry resources in FPGAs. Conventionally the ASIC implementation of online adders is based on the 4:2 compressors, as outlined in Fig.~\ref{Fig:PreviousWork} within the gray background. However, directly applying this approach in the FPGAs could be less efficient. This is because there is no carry propagation between the 2 FAs within a 4:2 compressor, and the net delay between them can be large. Instead, Kamp et al \cite{FPT09RA} and Hormigo et al \cite{ASAP09RA} described very similar mapping techniques for online adders with SD representations and CS representations, respectively. They share the common idea that maps the logic block (\texttt{LB}) within the dotted circle in Fig.~\ref{Fig:PreviousWork}. In this case, the fast-carry logic in the FPGA can be employed, and the delay between the 2 FAs can be largely reduced.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder_42comp_Spartan.pdf}
	\caption{Map the online adder onto Spartan FPGAs using the fast-carry resources. The grey background highlights the 4:2 compressor. Dotted circle indicates the logic block (LB) which can be mapped to the FPGA using the carry resources.}
	\label{Fig:PreviousWork}
\end{figure}

However, we notice the major limitations of both approaches that they only target on FPGAs with 4-input LUTs (4-LUT) and 2 LUTs within a logic Slice, such as the Xilinx Spartan series and the Altera Cyclone series. This is naturally reasonable because one \texttt{LB} can be mapped to a single Slice. Nevertheless, for FPGAs with 6-input LUTs (6-LUT) and 4 LUTs in a Slice, such as the Xilinx Virtex series and all Xilinx 7 series FPGAs, directly applying these approaches will result in either resource waste or logic fault. For instance, if two \texttt{LB}s are mapped to a slice with 4 LUTs as seen in Fig.~\ref{Fig:PreWorkVirtex6}, the outputs of \texttt{LB1} will be faulty because the its carry input cannot be explicitly initialized. Additionally, the output of \texttt{XORCY2} should not be used.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/SDadder_FastCarry_PreviousWork.pdf}
	\caption{An example illustrating that direct applying approaches in \cite{FPT09RA} on a Virtex-6 FPGA will results in faulty outputs, because the carry input of \texttt{LB1} cannot be explicitly initialized.}
	\label{Fig:PreWorkVirtex6}
    \vspace{-2ex}
\end{figure}

\subsection{Proposed Mapping Method}
To tackle this problem, we firstly modify the structure of the online adder to enable an efficient FPGA mapping. The structure of a 4-digit online adder is given as an example in Fig.~\ref{Fig:ModifiedOA}. In this equivalent structure, the first FA in each 4:2 compressor is split into 2 parts, which only generate carry and sum respectively. In this case they can be mapped individually on 2 LUTs. The second FA, which generates the outputs, is unchanged and can be implemented using the fast-carry logic.

@@

\begin{figure*}[tbp]
	\centering
	\includegraphics[width=.85\textwidth]{./Figures/SDadder_FPGA_New_4digit.pdf}
	\caption{Modified structure of online adder. Left: an example of 4-digit online adder. The shaded part refers to the 2 logic blocks (LBs) that can be mapped onto 1 slice. Right: one LB. The dotted box outlines the logic that can be mapped onto 1 LUT and the corresponding fast-carry logic.}
	\label{Fig:ModifiedOA}
\end{figure*}

The detailed slice mapping of the 2 \texttt{LB}s and the logic in each LUT are shown in Fig.~\ref{Fig:SliceNew}. The I/O signals are identical to the previous example in Fig.~\ref{Fig:ModifiedOA}. Normally 6-LUT can be configured with two different output ports \texttt{O6} and \texttt{O5}. For \texttt{LB1}, the carry input can be initialized by setting the \texttt{O6} of LUT2 to 0 constantly. In this case, the output of \texttt{MUXCY2} is always \texttt{O5} of LUT2, and the carry from \texttt{LB2} will not affect the results of \texttt{LB1}. For those logics to generate the MSD and the LSD, they can be combined into 1 Slice as observed from Fig.~\ref{Fig:ModifiedOA}. Therefore using this mapping method the resources within 1 slice can be fully utilized, potentially leads to significant area reduction.

@@

\subsection{Performance Analysis}\label{subsec:OA_Performance}
In our experiments, we compare the proposed online adder (\texttt{OA\_new}) against the original design which is implemented using behavioural descriptions based on 4:2 compressors (\texttt{OA\_behv}).  In addition, we also include ripple carry adder (\texttt{RCA}) which uses conventional arithmetic, into the comparisons. All designs are targeting on Xilinx Virtex-6 FPGA with speed grade -1. Notice that in order to utilize the carry resources and to avoid logic optimizations by the synthesis tool, the Xilinx primitive component "\texttt{Carry4}" is used to create \texttt{OA\_new}.

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.1. It can be seen that both designs with online arithmetic achieves relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay in the online adder is theoretically a fixed value and it is irrelevant to the operand precisions. In comparison, the rated frequency of \texttt{RCA} drops drastically for larger designs. In this case, both online adders outperform \texttt{RCA} for large word-lengths. Besides, our approach operates constantly faster than \texttt{OA\_behv}. The speed-ups of \texttt{OA\_new} over \texttt{OA\_behv} are from $20\%$ to $26\%$.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/Exp/Adder_Freq.eps}
	\caption{Rated frequencies of the RCA and the online adder with different implementation methods. The results are obtained from post place and route timing reports in ISE 14.1.}
	\label{Fig:AdderFreq}
	\vspace{-1ex}
\end{figure}

We also compare the resource usages of all three adder designs with respect to a variety of operand word-lengths. The results are presented in Fig.\ref{Fig:AdderArea}, in which the theoretical area budgets of the online adder predicted by (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}) are also plotted with dotted lines. Several observations can be made from the results. Firstly the predicted resource usages of both LUTs and Slices match well the results obtained from real implementations. Secondly it can be seen that the behavioural implementation \texttt{OA\_behv} is not area efficient for FPGAs, because the dedicated FPGA resources are not fully utilized. In comparison, our implementation \texttt{OA\_new} achieves significant area savings: $25\%\sim 33\%$ in LUTs and $67\%\sim77\%$ in Slices. Thirdly, our design still costs more area than \texttt{RCA}, with the area overhead of $80\%\sim98\%$ in terms of LUTs, and $0\%\sim88\%$ in terms of Slices. This is because for a single adder, the area overhead brought by the redundant number representations cannot be removed. 
%However research has shown that this overhead can be reduced for multiple-operand adders or adder trees [xxx].

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_LUTs2.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_Slices2.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different binary adder implementations with respect to a variety of operand word-lengths. (a) Usage of LUTs. (b) Usage of Slices.}
  \label{Fig:AdderArea}
  \vspace{-1ex}
\end{figure}

\section{Digit Parallel Online Multiplier on FPGAs}\label{Sec:OM_FPGA}
\subsection{Algorithm of Digit Parallel Online Multiplication}
Algorithm \ref{Algorithm:OnlineMult} as described in Section~\ref{Sec:Background} can be synthesized into a unrolled digit parallel structure. Recent study has shown that the digit-parallel online multiplier (\texttt{OM}) is more tolerant to timing violations in comparison to the conventional arithmetic, because timing errors initially affect the least significant digits and this will lead to small error magnitude~\cite{SKDAC14}. This implementation method is straightforward and can be directly utilized to design other digit parallel online operators.

On the other hand, significant area budget is required in order to implement this type of \texttt{OM}. Typically this is due to three main reasons. First, in comparison to the conventional multiplier, more iteration stages are needed. For instance in an $N$-digit \texttt{OM}, totally $(2N+\delta)$ iterations are required to generate $2N$ digits outputs, while $(N+\delta)$ iterations to generate the most significant $N$ digits result. Second, the logic of each stage is complex as it involves operations such as online addition, digit-vector multiplication, shifting and other combination logic blocks such as the selection function. Besides, the word-length of signals $H_{[j]}$, $W_{[j]}$ and $P_{[j]}$ is $(N+2+\delta)$, because the selection function needs 2 integer digits as well as 1 fraction digits to generate $z_j$, as described in Section~\ref{Sec:Background}. Whereas in an $N$-digit array multiplier, there are only $(N-1)$ rows, each of which is basically an $N$-digit ripple-carry adder. Third, this \texttt{OM} architecture is not optimized specifically for FPGA technology.

We address the first 2 design issues by providing an alternative online multiplication algorithm. Instead of simply implementing the unrolled format of Algorithm \ref{Algorithm:OnlineMult}, it can be optimized specifically for digit parallel operations. We notice that the online delay $\delta$ is employed in Algorithm~\ref{Algorithm:OnlineMult} because the input data are available in a digit serial fashion. Hence $\delta$ is used to accumulate enough input digits to generate the MSD of the result. Nevertheless if all digits of inputs are given simultaneously, $\delta$ is no longer necessary. Due to the same reason, $H[j]$, which takes one digit of each input per stage as shown in the original algorithm, can be optimized to take one partial product $Xy_j$ or $Yx_j$ per stage. To summarize, the modified algorithm for digit parallel online multiplication is described in Algorithm~\ref{Algorithm:OM_DigitParallel}, where $N$ denotes the operand word-length, function $frac(W_{[N]})$ refers to select the fractional digits of $W_{[N]}$, and the selection function $sel()$ is given previously in (\ref{Eq:SelFunc_OM}).
